Analysis of NLA17 Dissolved Gas Indicator

### 2021-03-09 scripts/dgIndicatorAnalysis.Rmd
Added text describing random semivariogram and did random semivariogram by ecoregion.

The water quality variables of pH, tnl, tpl, and chla showe strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.  The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation. The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation over the Euclidean distances that the lakes were sampled.  CH4 shows some autocorrelation to ~ 1000 km, but with no sill or leveling off the semivariance.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)
names(dg_1)
# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
dg_1 <- filter( dg_1, sample_source == "DG")

@@

>>>>

The water quality variables of pH, tnl, tpl, and chla show strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.  The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation. The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation over the Euclidean distances that the lakes were sampled.  CH4 shows some autocorrelation to ~ 1000 km, but with no sill or leveling off the semivariance.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)
names(dg_1)
# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData

dg_1 <- filter( dg_1, sample_source == "DG")
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")

@@

## Random Ecoregional Semivariogram N2O for ASLO Abstract
I used ecoregional-specific data frames to feed into random semivariogram function.  
```{r, ecoregion_semivar}

# unimodal, but not symmetrical
filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER") %>%  ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()

# zcpl <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="CPL")
# zcpl_variog <- variogram(log(dissolved_n2o_nmol) ~1, zcpl)
# plot(zcpl_variog, ylab = "N2O zCPL Semivariance")
# the 2-step process above matches the piped process below
xer_n2o_sf1 <- filter(dg_sf1,!is.na(dissolved_n2o_nmol)) %>%
  filter(ag_eco9 =="XER") %>% 
  variogram(log(dissolved_n2o_nmol) ~1, .)
plot(xer_n2o_sf1, ylab = "N2O XER Semivariance")

xer <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER")

print(xyplot(gamma ~ dist, wmt_n2o_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
    panel = function(x, y, ...) {
        for (i in 1:100) {
            xer$random = sample(xer$dissolved_n2o_nmol)
            v = variogram(log(random) ~ 1, xer)
            llines(v$dist, v$gamma, col = 'grey')
        }
        panel.xyplot(x, y, ...)
    },
    ylim = c(0, 0.35), xlab = 'distance', ylab = 'XER N2O semivariance'
))
